### 2021-10-28 plos_latex_template.tex
federico's and laura's edits plus many replies and edits responding to them, by rocío and andrea

\newcommand{\fede}[1]{\textcolor{ForestGreen}{Fede: #1}}
\newcommand{\rocio}[1]{\textcolor{MidnightBlue}{Rocio: #1}}
\newcommand{\doro}[1]{\textcolor{RedViolet}{Dorothea: #1}}
\newcommand{\jani}[1]{\textcolor{YellowOrange}{Janina: #1}}
\newcommand{\la}[1]{\textcolor{RawSienna}{Laura: #1}}

@@

The authors of this article participated in the organization of the annual user conference of the R Project for Statistical Computing, held in July 2021. useR! conferences are non-profit events organized by volunteers from the R community and managed by the R Foundation. The conference attracts a broad range of participants from academia, industry, government, and the non-profit sector. For 2021, we aimed to build a high-quality virtual and explicitly global conference in a kind, inclusive, accessible, and welcoming environment for everyone. 
In this article, we streamline our most important learnings in 10 simple rules to host an inclusive conference. These rules apply equally to academic, industry, or mixed conferences; the rules are inspired by a global experience but also apply at the regional or local level. The rules were learned during a virtual conference, but serve also as guidelines to in-person and hybrid events. \fede{A comment from Fede} \rocio{an ideal reply to that from Rocio} \doro{Doro chiming in, before the comment is resolved}

@@

% Rocío: Main message of the paragraph: Why we are writing this piece
Conferences, from the Latin \textit{conferre}--`to bring together', are spaces to meet and reconnect with members from a specific community, learn about advances in the field, and share our recent contributions.
However, conferences are likely to reproduce the systematic discrimination occurring in other spaces in our fields. 
Lack of representation and unwelcoming--or overtly aggressive--environments exclude people from the opportunities for learning, sharing, and producing knowledge, and becoming active community members, all of which are readily provided to non-marginalized people \cite{hendersonThoughtfulGatheringsGendering2020}.
Exclusionary conference experiences can divert career paths, affect lives, and drive people out of academia \cite{biggsAcademicConferenceChilly2018, hendersonThoughtfulGatheringsGendering2020}. 
This message is not new. 
Barriers such as unattainable registration costs, sexism, and ableism, among others, have already been exposed and discussed in the literature \cite{arendDisparityConferenceRegistration2019, biggsAcademicConferenceChilly2018, depickerRethinkingInclusionDisability2020a, irishIncreasingParticipationUsing2020}, 
and some proposals for more inclusive conferences have been put into practice \cite{gichoraTenSimpleRules2010a, levitisCenteringInclusivityDesign2021, atkinsonJournalMedicine20202021, foramittiVirtuesVirtualConferences2021, ninerBetterWhomLeveling2021, rabyMovingAcademicConferences2021}, with a primary focus on the online format to open doors for inclusion. %https://www.cscce.org/resources/organizing-community-events/

% Rocío: Main message of the paragraph: Roadmap of the paper
This article puts the focus on practices for inclusive conferences, whether they are virtual or in-person, though we advocate for having a strong online component, well-articulated with the in-person component, for all conferences.
The rules written here are directed to people who are part of a stable meetings committee that oversees the site/location selection process, or that coordinates with the local organizers of conferences.
The rules can also be helpful to local/virtual organizers who desire to make an inclusive conference starting at the planning stage.
These tips stem from the authors' experience of organizing useR! 2021, a virtual and global statistical computing conference for users and developers of the R programming language \cite{r_core_team_2021}. 
We embraced the challenge of organizing a high-quality virtual conference in the context of the COVID-19 pandemic and making it a kind, inclusive, and accessible experience for as many persons as possible. 
Here, we share the lessons learned within the past year of organizing useR! 2021, summarized as ten simple rules towards hosting an inclusive conference.
The rules are organized in three groups (Figure \ref{fig:diagram}). %original source: https://docs.google.com/drawings/d/1iS1pLc9OldLMe_jhNT6OGIbT9s5KvlHpUoDd19vHkCY/edit?usp=sharing
%sara: still need to better format figure size and get the alt text right :P
Group 1 includes rules 1, 2, and 10, which refer to pillars of an inclusive conference: embracing diversity in all its dimensions, creating a safe and welcoming environment for everyone, and making the conference part of a long-term process for inclusion.
The second group includes rules 3 and 4 and focuses on the people who participate in the conference. 
Rule 3 refers to the importance of working with an inclusive and diverse organizing team, and Rule 4 concerns the necessity of removing implicit and systemic bias from spotlight roles like keynote speakers, other presenters, program committee members, or other session chairs. 
The third group includes rules 5 to 9. These rules are about components of the conference that should be carefully planned for: an online component, accessibility to people with disabilities, language inclusiveness, a welcoming communication strategy, and financial resources to support inclusion. 
These rules apply equally to academic, industry, or mixed conferences. While these rules are inspired in a global virtual experience they also apply at the regional or local level and can inform in-person and hybrid events.

@@

% keeping references:
%   However, conferences can become discriminating spaces, in which members of some specific privileged groups reproduce the systematic inequalities that occur in academia and society \cite{arendDisparityConferenceRegistration2019, timperleyHeMoanaPukepuke2020, gewinWhatScientistsShould2019, brownAbleismAcademiaWhere2018}.
% Liz adding reference marks2021meeting

@@

% Rocío: Main idea: what is diversity and why does it matter? (inequalities)
Diversity encompasses multiple dimensions: age, physical ability, career stage, gender, gender identity, geographic origin, language, neurodiversity, race, religion, sexual orientation, and socioeconomic background, to name a few.
Human diversity should be celebrated and respected in every way. Nonetheless, we live in a world with implicit hierarchies along these axes. Some statuses (e.g., cisgender, white, male, from the United States or Western Europe) hold the privilege of being the defaults centered around which all systems--including conferences--are consciously and unconsciously built.  
While no isolated initiative can change reality by itself, building a more diverse and inclusive conference starts by recognizing that these inequalities have systematically excluded whole groups of people from academia and scientific and professional circles \cite{timperleyHeMoanaPukepuke2020}.

% Rocío: more effort towards including the more excluded
Recognizing our privileges--unearned advantages given by society to some people but not all--particularly in our field and in our scientific or professional community, will help identify which subgroups have been the most excluded or discriminated against. 
These are the groups we need to make more effort to include. 
Investing more effort in the most excluded groups does not mean neglecting the others, but it will guide the vision of diversity for your conference--and your strategies to achieve it.
Will a more diverse conference translate into a gender distribution of your speakers that is representative of the general population? 
Would it be the presence of racialized people --especially Black people-- among the head organizers, speakers, and attendees? 
Would it be having LGBTQIA+ friendly-spaces or community participation from key geographic regions?
%"The answer to these questions depend on your field, region or community.
Intersectionality, or the complex, cumulative way in which the effects of multiple forms of discrimination combine or intersect, should be taken into account.

@@

% Rocío: main idea: Inclusion for everyone
While it is essential to improve representation towards some of the most visible dimensions of human diversity, such as race, gender, and country of origin, building a truly inclusive environment means taking care of all the other aspects of diversity as well. 
Having consideration of religious practices, setting specific accommodations for breastfeeding and child care, having LGBTQIA+-friendly spaces, creating community-only spaces, and acknowledging that gender is not binary are just some examples of decisions that can make inclusion real.
Importantly, you can take active steps in creating a more welcoming environment without requiring anyone to disclose personal information.

% Rocío: main idea: Code of conduct for a safe place
Adopting a code of conduct and gathering a team to enforce it are key aspects in creating a safe environment during a conference \cite{favaroYourScienceConference2016}.
%Dorothea: it is not a document, right? The document will not keep the community safe. Maybe say "a set of rules" in case someone doesn't know what a code of conduct is?
The code of conduct is a document meant to keep the community safe and should state clearly the unacceptable behaviors, the consequences for engaging in such behavior, and the way to report violations \cite{auroraHowRespondCode2019}. 
The code of conduct team must receive training on how to receive reports, respond to incidents, and communicate their responses, and organize accordingly. 
A diverse code of conduct team will be more understanding of power dynamics and sensitive to discrimination and harassment issues. 
There are also disciplinary cultures and geographical considerations that may need to be considered when developing and communicating the code of conduct. 
We strongly recommend reading `How to Respond to Code of Conduct Reports' \cite{auroraHowRespondCode2019} as an excellent starting point for the code of conduct teamwork.
% Rocío: two of these sentences, diverse CoC team, and communication of CoC, may belong to rules organizing team and communication, respectively.

% Rocío: Could have a paragraph about active bystandership—Andrea
%ast: tiny: check CoC capitalization (LA)

@@

% Rocío: Idea of this paragraph: build a diverse team
A genuinely inclusive conference can only be organized by an inclusive and diverse organizing team, in charge of the conference decision-making.
Build a team with people from different regions, genders, races and ethnicities, socioeconomic statuses, and other aspects of diversity. 
Go beyond balancing all genders in this effort and pay special attention to to intersectional diversity (see \textbf{Rule \ref{rule_diversity}}). 
To ensure a deep understanding and smooth communication with different diverse groups, it is essential to create a representative working group that functions as a snapshot of the general population.
If you already have an organizing team, check for gaps in its composition.

% Rocío: Idea of this paragraph: real inclusion when having diverse members
Gathering a diverse team will only work if there is real inclusion. 
People with disabilities often say: `Nothing about us without us'; the same holds for other dimensions of diversity. 
This means that the actual life experiences, expertise, and insights from people in marginalized groups cannot be replaced by good intentions from people outside these groups \cite{costanzachockDesign2020}.
People who have experience with exclusion have the social and technical expertise about removing barriers to inclusion.
A truly inclusive and welcoming space is one in which everyone in the team is invited to be part of the solution.
However, do not expect people from minoritized groups to only work on these solutions, and respect their interests in other areas of the organization.

% Rocío: Idea of this paragraph: positive outcomes of having a diverse team - keeping it general at least
Creating and maintaining such a team and space may seem more challenging than working in homogeneous teams--when you are not used to them--but the positive outcomes are worthwhile. 
Having diverse people in decision-making positions will affect positively all the other aspects of your conference because all the processes will benefit from their input, expertise, and distinct perspectives \cite{hongGroupsDiverseProblem2004}. 
In addition, a diverse team plays an important role at creating a welcoming space because representation--seeing people with similar life experiences occupy public spaces, positions of power, and breaking negative stereotypes--is one of the best ways to create a sense of belonging for everyone participating in the conference (see \textbf{Rule \ref{rule_inclusion}}).

% Rocío: Idea of this paragraph: tips to make a diverse and inclusive team work.
The following advice should apply to any kind of team, but is specially relevant when working with people from marginalized groups. 
Do not expect self-nomination and voting to work as mechanisms to counteract systemic inequalities. Instead, nominate directly and offer leading positions, and let people from privileged groups step down.
Build an environment in which every person can express their position and give priority to people from systematically excluded groups.
Offer support and encouragement if you encounter cases of impostor syndrome. 
You can, for example, break the expectations about leadership as a lonely task and create smaller, co-led groups, where everyone finds their preferred tasks and gets to take leadership.

% Rocío: Idea of this paragraph: take care of your team.
Most importantly, take care of your team. 
Having a diverse team and executing inclusive and accessible practices throughout the organizing period--it might be a year or two--may require a lot of effort. 
%Dorothea: I would remove the following sentence, I think we have made the point that it is worthwhile
The effort is worth it because it strengthens the event and the community making it truly welcoming for everyone. 
However, having a strong community as the only reward may be enough for those who are in more privileged positions. 
Some people and often the minoritized ones do not have institutional support to put time and effort into the organization tasks and do not have the luxury to commit to the organization for free; consider paying them as an item in your budget.  
In addition, tasks such as receiving and responding to code of conduct reports, can be emotionally intense work and should be additionally rewarded.
Prioritize your team's well-being. Check on them regularly and make sure everyone is comfortable. 
Be mindful of each particular context, be flexible with hours or commitments, and revise your budget (see \textbf{Rule \ref{rule_financial}}).

% % Rocío: might go to the next rule
% Splitting the workload and responsibilities should not be done by putting care-taking labors--community building, meeting organization and note-taking, conversations with potential partners--on the hands of women and other minoritized groups, while people from privileged groups take the lead in stereotypical highly-valued tasks (see \textbf{Rule \ref{rule_unbias}}). Make work visible, discuss collectively what should be done (and why), divide and account work. Being transparent and sharing the information is key to make work visible and collective. %sara: this relates to make labor visible from DF, maybe we can cite them here too. %Rocío: sure.

%This is specially important for networking spaces, where receiving ideas, direct feedback, and spaces to co-lead from a wide variety of community members can be the key to organize meaningful sessions for all the conferences attendees. This gave a sense of variety and an appeal of "at least one networking session for everyone".  (Marcela)
%ast 17sep: i think we can talk here about planning neworking here, but also: the whole paragraph applies to team dynamics in general. Sara: maybe you can use this as we spoke relating team work here. And I think the working together part is not here but the following paragraph 
%sara: The last sentence in the paragraph sounds odd to me, but maybe I'm missing something. Andrea, I added a few sentences to the following paragraph regarding work dynamics
% Rocío: the rule is getting too long and, since this is focused on networking, I'd prefer it to go to a networking section.

\section{Consciously mitigate bias in your spotlight roles}
\label{rule_unbias}

%idea of paragraph 1: our lists are biased
When choosing or inviting people as keynote speakers, program committee, session chairs, and other spotlight roles, it is likely that there will not be much diversity in the first set of names. 
Many of our biased lists are products of the existing systems that have always privileged some groups of people \cite{dwyerNoticeWhoScience2021,swartzScienceValueDiversity2019,wongBuildDiversityScience2020,dignazioUnicornsJanitorsNinjas2020}. 
Rather than deter us, this implicit and systemic bias should encourage us to look further to find great people that are not routinely in the spotlight. 
Ensuring diversity in each of these roles needs to be a deliberate process. 
We need to go beyond our narrow and often limited networks to look for, reach out to, invite, encourage, and onboard these people until there is ample representation across the diversity spectrum and dimensions. 
\doro{The last sentence sounds odd. Maybe just: "until there is ample representation."}
Make sure that every selection committee--the committee looking for keynote speakers, the selection committee for abstracts, the prizes and award committees--are also diverse \cite{swartzScienceValueDiversity2019, wongBuildDiversityScience2020}, and ask them to be aware that everyone has implicit biases, to recognize them, and try to counteract them. 
An inclusive and diverse organizing team (\textbf{Rule \ref{rule_organizing_team}}) is already a great starting point to overcome this bias in other roles. The regional and local communities in your field are also good sources to tap into.

% Rocío: main idea: unbias roles and give spotlight to the roles--and people who have contributed in roles--that are not stereotypically categorized as success
In academia and the industry, not only is the list of people considered in spotlight roles usually biased, but the spotlight roles themselves are.
Scientific publication and development of software, are often regarded with higher value than community building or teaching, which may be equally if not more challenging, less rewarded, and usually tackled by women, people of color, people with disabilities and from other minoritized groups \cite{cheng2020x+, burfordHomelinessMeantHaving2020}.
Defy the stereotypical criteria for success by acknowledging these community practices and the people behind them.
You can, for instance, give more space in your conference to talks about community building as an area of knowledge production, and reframe \doro{amend?} the awards ceremony to acknowledge those who prepared accessible slides and presentations (see Rule \ref{rule_accessibility}), for being mindful of inclusiveness. \doro{Suggestion: "amend the awards ceremony to also acknowledge those who prepared accessible slides and presentations, contributed to community building or presented teaching materials"}
Finally, do not restrict people from marginalized groups or community-builders to talk or work only on issues related to diversity and inclusion; recognize their areas of expertise and interest and offer them space to showcase them.

\section{Have a strong online component of the conference} 
\label{rule_online}

% Rocío: main idea of the paragraph: online conferences can open opportunities for inclusion
In-person interaction in conferences is priceless, but only for the ones who can afford to attend. 
Barriers such as cost of registration, transport and accommodation, the logistics of long-distance travel, and discriminatory visa applications, are particularly true for conferences that usually take place in high-income countries \cite{arendDisparityConferenceRegistration2019,gewinWhatScientistsShould2019,jooKeepOnlineOption2021}. 
Online conferences are more inclusive because they do not require a visa or a big budget, and are more accessible to people who may be unable to travel because of health issues or family responsibilities \cite{salibaGettingGripsOnline2020}.
This means that online conferences have a greater reach, not only in terms of participants but in terms of the tutors and presenters that can participate \cite{atkinsonJournalMedicine20202021, roosOnlineConferencesNew2020, jooKeepOnlineOption2021}.
The online format may also make it easier to be inclusive of geographic regions by encompassing several timezones in the week or easily deciding to rotate the favored timezone year to year without depending on a conference central location.

% Rocío: main idea of the paragraph: In the context of reopening, hybrid conferences may be a thing. They also need to be inclusive.
Within the context of the COVID-19 pandemic, many conferences embraced the online format; but at the time of writing, some are reverting to in-person, which risks going back to the barriers mentioned above.
Alternatively, a conference could have a hybrid format with an in-person and an online components that articulate well into a hybrid format that maximizes the experience of everyone attending your conference. This dual format could allow a group of people to interact face-to-face while providing many others the opportunity to participate remotely. 
The challenge and requirement for this kind of setting would be to make the online component as relevant as the in-person component and not just a consolation prize to the less privileged in the community \cite{ninerBetterWhomLeveling2021}.
%Needless to say, an online component is not inclusive \textit{per se} without taking many other precautions, and it does not eliminate the need to organize an inclusive in-person component. 
%For instance, to avoid favoring the same regions in every edition, we suggest rotating the location of the conference—if it is global, moving from one continent to another. 
% Rocío: Now I'm not sure if we need the last two sentences that I commented.

>>>>

\newcommand{\fede}[1]{\textcolor{ForestGreen}{Fede: #1}}
\newcommand{\rocio}[1]{\textcolor{MidnightBlue}{Rocio: #1}}
\newcommand{\doro}[1]{\textcolor{RedViolet}{Dorothea: #1}}
\newcommand{\jani}[1]{\textcolor{YellowOrange}{Janina: #1}}
\newcommand{\la}[1]{\textcolor{RawSienna}{Laura: #1}}
\newcommand{\as}[1]{\textcolor{Violet}{Andrea: #1}}

@@

The authors of this article participated in the organization of the annual conference of the R Project for Statistical Computing, held in July 2021. useR! conferences are non-profit events organized by volunteers from the R community and managed by the R Foundation. The conference attracts a broad range of participants from academia, industry, government, and the non-profit sector. For 2021, we aimed to build a high-quality virtual and explicitly global conference in a kind, inclusive, accessible, and welcoming environment for everyone. 
In this article, we outline our most important lessons learned in 10 simple rules to host an inclusive conference.\fede{general comment on this: we should clearly state that this whole-virtual setting was a clear consequence of the pandemics, AND it was planned like this from the beginning, and not changing in-between to this format} These rules apply equally to academic, industry, or mixed conferences; the rules are inspired by a global experience but also apply at the regional or local level. The rules were learned during a virtual conference, but serve also as guidelines to in-person and hybrid events.

@@

% Rocío: Main message of the paragraph: Why we are writing this piece
Conferences, from the Latin \textit{conferre}--`to bring together', are spaces to meet and reconnect with members from a specific community, learn about advances in the field, and share our recent contributions.\fede{between this and the next sentence, there is a strong break. need some ligation, something like "Despite the many efforts in this direction, ..."}
However, conferences are likely to reproduce the systematic discrimination occurring in other spaces in our fields. 
Lack of representation and unwelcoming--or overtly aggressive--environments exclude people from opportunities for learning, sharing, and producing knowledge, and becoming active community members, all of which are readily provided to non-marginalized people \cite{hendersonThoughtfulGatheringsGendering2020}.
Exclusionary conference experiences can divert career paths, affect lives, and drive people out of academia \cite{biggsAcademicConferenceChilly2018, hendersonThoughtfulGatheringsGendering2020}. 
This message is not new. 
Barriers such as unattainable registration costs, sexism, \as{prejudice against disable people ( ableism)}, among others, have already been exposed and discussed in the literature \cite{arendDisparityConferenceRegistration2019, biggsAcademicConferenceChilly2018, depickerRethinkingInclusionDisability2020a, irishIncreasingParticipationUsing2020}, 
and some proposals for more inclusive conferences have been put into practice \cite{gichoraTenSimpleRules2010a, levitisCenteringInclusivityDesign2021, atkinsonJournalMedicine20202021, foramittiVirtuesVirtualConferences2021, ninerBetterWhomLeveling2021, rabyMovingAcademicConferences2021, noauthor_discover2021}, with a primary focus on the online format to open doors for inclusion.\fede{do we need maybe a short definition of the concepts of diversity-equity-inclusion for us/refer to something?} %LA: https://www.cscce.org/resources/organizing-community-events/

% Rocío: Main message of the paragraph: Roadmap of the paper
This article focuses on practices for hosting inclusive conferences.
The rules written here are directed to people who are part of a stable meetings committee that oversees the site/location selection process, or that coordinates with the local organizers of conferences.
The rules can also be helpful to local/virtual organizers who intend to put inclusion at the core of the planning phase.
These tips stem from the authors' experience of organizing useR! 2021, a virtual and global statistical computing conference for users and developers of the R programming language \cite{r_core_team_2021}. 
We embraced the challenge of organizing a high-quality virtual conference in the context of the COVID-19 pandemic and making it a kind, inclusive, and accessible experience for as many persons as possible. 
Here, we share the lessons learned within the past year of organizing useR! 2021, summarized as ten simple rules towards hosting an inclusive conference.
The rules are organized in three groups (Figure \ref{fig:diagram}). %original source: https://docs.google.com/drawings/d/1iS1pLc9OldLMe_jhNT6OGIbT9s5KvlHpUoDd19vHkCY/edit?usp=sharing
%sara: still need to better format figure size and get the alt text right :P
Group 1 includes rules 1, 2, and 10, which refer to pillars of an inclusive conference: embracing diversity in all its dimensions, creating a safe and welcoming environment for everyone, and making the conference part of a long-term process for inclusion.
The second group includes rules 3 and 4 and focuses on the people who participate in the conference. 
Rule 3 refers to the importance of working with an inclusive and diverse organizing team, and Rule 4 concerns the necessity of counteracting implicit and systemic bias from spotlight roles like keynote speakers, other presenters, program committee members, or other session chairs. 
The third group includes rules 5 to 9. These rules are about components of the conference that should be carefully planned for: an online component, accessibility for people with disabilities, language inclusiveness, a welcoming communication strategy, and financial resources to support inclusion. 
These rules apply equally to conferences where the expected audience is academic, from the industry, or a mixed group. While these rules are inspired in a global virtual experience they also apply at the regional or local level and can inform in-person and hybrid events.
\fede{maybe something like: "Even if this set of rules is derived from our experience with a global scale event, we believe these can be easily transferred to events with regional or local scopes"}

@@

% keeping references:
%   However, conferences can become discriminating spaces, in which members of some specific privileged groups reproduce the systematic inequalities that occur in academia and society \cite{arendDisparityConferenceRegistration2019, timperleyHeMoanaPukepuke2020, gewinWhatScientistsShould2019, brownAbleismAcademiaWhere2018, marks2021meeting}.

@@

% Rocío: Main idea: what is diversity and why does it matter? (inequalities)
Diversity encompasses multiple dimensions: age, physical ability, career stage, gender, gender identity, geographic origin, language, neurodiversity, race, religion, sexual orientation, and socioeconomic background, to name a few.
Human diversity should be celebrated and respected in every way. Nonetheless, we live in a world with implicit hierarchies along these axes. Some statuses (e.g., cisgender, white, male, from the United States or Western Europe) hold the privilege of being the defaults around which all systems--including conferences--are consciously and unconsciously built. \as{People outside these groups suffer different forms of oppression, like sexism, racism, ableism, homophobia, and transphobia. Furthermore, no diversity dimension operates in isolation from the others, they intersect in complex ways to determine people's experience in the world--a concept know as \textit{intersectionality}--so that anyone can suffer from some forms of oppression and be privileged along some other dimension, and some groups are particularly vulnerable and systematically left behind.}
While no isolated initiative can change reality by itself, building a more diverse and inclusive conference starts by recognizing that these inequalities have systematically excluded whole groups of people from academia and scientific and professional circles \cite{timperleyHeMoanaPukepuke2020}.

% Rocío: more effort towards including the more excluded
Recognizing our privileges--unearned advantages given by society to some people but not all--particularly in our field and in our scientific or professional community, will help identify which subgroups have been the most excluded or discriminated against. 
These are the groups we need to make more effort to include. \fede{General thought: I like the way we say WE and YOU. I still think I saw most such 10SR papers speak to the readers addressing them with the roles. Which would be "Organizers", "Attendees". Just my 2 cents as avid reader of 10SR papers ;)}
Investing more effort towards the most excluded groups does not mean neglecting the others, but it will guide the vision of diversity for your conference--and your strategies to achieve it.
Will a more diverse conference translate into a gender distribution of your speakers that is representative of the general population? 
Would it be the presence of racialized people --especially Black people-- among the head organizers, speakers, and attendees? 
Would it be having LGBTQIA+ friendly-spaces or community participation from key geographic regions?
The answer to these questions depend on your field, region, or community.

@@

% Rocío: main idea: Inclusion for everyone
While it is essential to improve representation towards some of the most visible dimensions of human diversity, such as race, gender, and country of origin, building a truly inclusive environment means taking care of all the other aspects of diversity as well. 
Accommodating for religious practices, setting specific accommodations for breastfeeding and child care, scheduling events exclusive for diverse groups (e.g., LGBTQIA+-friendly spaces), promoting and respecting pronouns for all genders and having gender-neutral bathrooms, are just some examples of decisions that are relatively easy to implement and can have a large impact in making inclusion real (see \cite{noauthor_discover2021} for some great advice on `low-hanging fruits').
Importantly, by being proactive in creating such a welcoming space, you can be respectful of everybody's privacy and avoid requiring anyone to disclose personal information–such as revealing a disability, the sexual orientation, gender identity, or mental health issues, just to a name a few.

% Rocío: main idea: Code of conduct for a safe place
Adopting a Code of Conduct and gathering a team to enforce it are key aspects in creating a safe environment during a conference \cite{favaroYourScienceConference2016}.
The Code of Conduct is a document meant to keep the community safe and should state clearly the unacceptable behaviors, the consequences for engaging in such behavior, and the way to report violations \cite{auroraHowRespondCode2019}. 
The Code of Conduct team must receive training on how to receive reports, respond to incidents, communicate their responses, and organize accordingly. 
Since people from marginalized groups are more likely to be victims of Code of Conduct violations and to see their claims dismissed due to power dynamics, assembling a diverse Code of Conduct team will be essential to deal with discrimination and harassment issues and the power dynamics underneath them and deal with them in a sensitive, victim-centered manner. It may also be more understanding of different disciplinary cultures and geographical and cultural considerations. 
We strongly recommend reading `How to Respond to Code of Conduct Reports' \cite{auroraHowRespondCode2019} as an excellent starting point for the Code of Conduct teamwork.
% Rocío: two of these sentences, diverse CoC team, and communication of CoC, may belong to rules organizing team and communication, respectively.

% Rocío: Could have a paragraph about active bystandership—Andrea%ö

@@

% Rocío: Idea of this paragraph: build a diverse team, and representation

A genuinely inclusive conference can only be organized by an inclusive and diverse organizing team, in charge of the conference decision-making.
Organizers should build a team with people from different regions, genders, races and ethnicities, socioeconomic statuses, and other aspects of diversity.
Having diverse people in decision-making positions will affect positively all the other aspects of your conference because all the processes will benefit from their input, expertise, and distinct perspectives \cite{hongGroupsDiverseProblem2004}. 
In addition, a diverse team plays an important role at creating a welcoming space because representation--seeing people with similar life experiences occupy public spaces, positions of power, and breaking negative stereotypes--is one of the best ways to create a sense of belonging for everyone participating in the conference (see \textbf{Rule \ref{rule_inclusion}}). If you already started assembling an organizing team, check for gaps in its composition. 
The regional and local communities, groups, or associations in your field are good sources to tap into.

% Rocío: Idea of this paragraph: non-transferability and do not restrict people
People with disabilities often say: `Nothing about us without us'; the same holds for other dimensions of diversity. This means that the actual life experiences, expertise, and insights from people in marginalized groups cannot be replaced by good intentions from people outside these groups \cite{costanzachockDesign2020}.
People who have experienced exclusion have the social and technical expertise to fight against it and they will be essential in any team.
However, they must have the freedom to choose in which areas of the conference they want to work and not be restricted to diversity and inclusion aspects.

% Rocío: Idea of this paragraph: tips to make a diverse and inclusive team work + paying them
When working with people from marginalized groups do not assume that self-nomination and voting will work as mechanisms to choose leadership positions. Instead, nominate directly and offer leading positions that would normally be occupied by people from privileged groups.
Build an environment in which every person can express their position and give priority to people from systematically excluded groups.
Many people lack the institutional support to put time and effort into the organization tasks and do not have the luxury to commit to the organization for free; consider paying them as an item in your budget.  
In addition, tasks such as receiving and responding to Code of Conduct reports can be emotionally intense work and should be additionally rewarded.

% % Rocío: might go to the next rule
% Splitting the workload and responsibilities should not be done by putting care-taking labors--community building, meeting organization and note-taking, conversations with potential partners--on the hands of women and other minoritized groups, while people from privileged groups take the lead in stereotypical highly-valued tasks (see \textbf{Rule \ref{rule_unbias}}).

\section{Consciously counteract bias in your spotlight roles}
% in the conference program
\label{rule_unbias}

%idea of paragraph 1: our lists are biased

When choosing or inviting people for visible and valued roles in the conference––such as keynote speakers, program committee members, session chairs––, it is likely that there will not be much diversity in the first set of names.
Our biased lists are products of the existing systems that have always privileged some groups of people \cite{dwyerNoticeWhoScience2021,swartzScienceValueDiversity2019,wongBuildDiversityScience2020,dignazioUnicornsJanitorsNinjas2020}. 
Rather than deter us, this should encourage us to go beyond our narrow and often limited networks to look for, reach out to, invite, encourage, and onboard great people that are not routinely in the spotlight, making sure that these roles do not stay with people in privileged groups (e.g. avoid `manels' and all-white panels, for instance).

Having a chance to present at a conference is an opportunity to gain--or keep--visibility in the community or field. 
Even diverse selection committees can be subjected to unconscious bias, by interpreting information such as the location of the institution or the names of the authors and attaching value to their work according to negative or positive stereotypes. 
Selection committees need to be reminded of such biases, and try to counteract them when evaluating submissions and presentations, which should have clear evaluation criteria.
\cite{swartzScienceValueDiversity2019, wongBuildDiversityScience2020}

% Rocío: main idea: unbias roles and give spotlight to the roles--and people who have contributed in roles--that are not stereotypically categorized as success

Activities such as scientific publication and development of software are often regarded as higher value than care-taking roles like community building or teaching. 
These roles may be equally if not more challenging and are usually assumed by women, people of color, people with disabilities, and other minoritized groups \cite{cheng2020x+, burfordHomelinessMeantHaving2020}. Research about these topics is equally undervalued compared to other areas of knowledge production. 
Counteracting bias also means broadening the range of topics of the conference and giving spotlight/visibility to usually undervalued research and the people behind it.
%Defy the stereotypical criteria for merit by acknowledging these community practices and the people behind them.
You can also reframe the awards ceremony to acknowledge those who contributed to community building, presented teaching materials, and assumed other care-taking roles.

\section{Have a strong online component} 
\label{rule_online}

% Rocío: main idea of the paragraph: online conferences can open opportunities for inclusion
In-person interaction at conferences is priceless but often this is true and possible only for the ones who can afford to attend. 
Barriers such as cost of registration, transport and accommodation, the logistics of long-distance travel, and discriminatory visa applications, are particularly true for conferences that usually take place in high-income countries \cite{arendDisparityConferenceRegistration2019,gewinWhatScientistsShould2019,jooKeepOnlineOption2021}. 
Online conferences are more inclusive because they do not require a visa or a big budget, and are more accessible to people who may be unable to travel because of health issues or family responsibilities \cite{salibaGettingGripsOnline2020}.
This means that online conferences have a greater reach, not only in terms of participants but in terms of the tutors and presenters that can participate \cite{atkinsonJournalMedicine20202021, roosOnlineConferencesNew2020, jooKeepOnlineOption2021}.
The online format may also make it easier to be inclusive of geographic regions by encompassing several timezones or deciding to rotate the favored timezone every year, without depending on a conference central location.

% Rocío: main idea of the paragraph: In the context of reopening, hybrid conferences may be a thing. They also need to be inclusive.
Within the context of the COVID-19 pandemic, many conferences embraced the online format; but at the time of writing, some are reverting to in-person, which risks going back to the barriers mentioned above \cite{jooKeepOnlineOption2021}.
Alternatively, a conference could have a hybrid format with an in-person and an online components \la{that articulate well into a hybrid format that maximizes the experience of everyone attending your conference}. This dual format could allow a group of people to interact face-to-face while providing many others the opportunity to participate remotely. 
The challenge and requirement for this kind of setting would be to make the online component as relevant as the in-person component and not just a consolation prize to the less privileged in the community \cite{ninerBetterWhomLeveling2021}.

%an online component is not inclusive \textit{per se} without taking many other precautions

%For instance, to avoid favoring the same regions in every edition, we suggest rotating the location of the conference—if it is global, moving from one continent to another. 
% Rocío: Now I'm not sure if we need the last two sentences that I commented.

% Rocío: main idea: Networking may seem the weakest part of online conference, but it doesn't have to be. 
% Rocío: I'm still doubting if this should be the 2nd paragraph, and I think that it needs some editing.
Networking and socializing have been mentioned as challenging aspects of online conferences, mostly because we are used to interactions at in-person social settings such as coffee or lunch breaks \cite{salibaGettingGripsOnline2020, roosOnlineConferencesNew2020}. 
However, in-person settings might not be comfortable or appealing to everyone for socializing and 
some of these spaces may be exclusionary (e.g. galas or dinner nights at expensive venues). 
On the other hand, there is evidence that virtual communication before face-to-face interaction can make people from minoritized groups feel more included, thus participate more (e.g. \cite{trianaDoesOrderFacetoFace2012, blackEngenderingBelongingThoughtful2020}). 
Organizers of online conferences should invest time in creating opportunities to meet and bond virtually, respecting people's limits, preferences, and remembering that `the usual' does not necessarily work for everyone, and that no single networking activity will ever serve the whole community. 
It is worth trying new and varied activities centered around different groups of people outside the mainstream of your conference %LA cita 
.
Some ideas in this line are: offering the option of written chat instead of voice or video conversations, opening events with teamwork like trivia, offer some events that can be enjoyed passively like movies, yoga sessions, or art displays, where attendants can choose to just sit and enjoy without talking, or have a chat channel to comment on their experiences during the session.\fede{another activity: virtual city tours, I had it as a corollary activity for a summer school} \la{Some more specifics on how to achieve these suggestions in this paper: https://osf.io/k3bfn/}

% we need to check if something here (newbies!) is not missing in the above paragraphs
% As we mentioned previously, conference attendees should have ample opportunities to network with each other. However, having a diverse offer of networking opportunities that can appeal to people with different backgrounds, accessibility needs, and preferences can be challenging, especially if you have the mindset of organizing each activity for the complete pool of attendants. If you think of smaller activities that can reach specific groups, preferably co-led by community leaders, these sessions can be more productive successful, and inclusive than trying to organize one single activity that can please the whole community (\textbf{Rule \ref{rule_unbias}}). Some examples are the newbies sessions for first-timers, mixers lead by specific subgroups or communities, or leisure activities that reunite subgroups with the same interests: arts, exercise, sports, movies, etc.
%Adithi had said: "Assist newbies of the conference navigate through the conference as they might feel overwhelmed attending a conference." but i see batool included newbies here :)
